---
title: |
  Exploiting Crowd-Compression Bias in a Proper-Scoring-Rule Prediction Tournament:
  A Statistical Forecasting System for the 2026 FIFA World Cup
author: "Souparneya Chakrabarti"
date: "July 11, 2026"
output:
  pdf_document:
    latex_engine: xelatex
    number_sections: true
    fig_caption: true
    toc: true
    toc_depth: 3
    citation_package: natbib
geometry: "margin=1.25in"
fontsize: 11pt
linestretch: 1.4
bibliography: references.bib
biblio-style: apalike
link-citations: true
header-includes:
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{float}
  - \usepackage{hyperref}
  - \usepackage{xcolor}
  - \usepackage{caption}
  - \usepackage{enumitem}
  - \usepackage{mdframed}
  - \hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue}
  - \setlength{\parskip}{6pt}
abstract: |
  This paper documents a solo research campaign conducted within the Jump Probability
  Cup (JTC), a 2026 FIFA World Cup prediction tournament on \textit{play.sportspredict.com}
  scored by Relative Brier Points (RBP)---a metric that rewards a submitted probability
  when it outperforms the platform's crowd consensus on the same question and penalises
  it when it does not.

  I built a complete statistical forecasting pipeline: point-in-time Elo ratings over
  49,473 historical international matches (1872--present), a Poisson goal-rate
  regression with a Dixon--Coles low-score correction and Negative-Binomial
  overdispersion correction, a parallel ordered-logit match-result model following
  \citet{hvattum2010}, and a StatsBomb event-level panel spanning the 2018 and 2022
  World Cups. The methodology is grounded in a systematic review of thirteen sources
  spanning goal-based modelling, Elo-based outcome regression, small-sample
  model-selection theory, and wisdom-of-crowds aggregation.

  Two findings anchor the paper. The first is operational: the JTC crowd consensus is
  structurally compressed toward 50\% relative to a calibrated estimate, characterised
  by $\hat{p}_{\text{crowd}} \approx 0.504 + 0.61 \times (\hat{p}_{\text{mine}} - 0.5)$,
  fit on 772 submitted questions across the full campaign ($r = 0.83$, 95\% CI on slope
  $[0.58, 0.64]$)---a relationship that holds essentially unchanged from an earlier fit
  at less than half the sample. The second is that four independent tests---a
  Platt-scaling recalibration, a logistic-regression/gradient-boosting blend evaluated
  under two out-of-sample validation schemes, a linear regression of RBP against
  crowd-deviation, and a closed-form result on the expected cost of pushing past a
  calibrated estimate under this scoring rule---all reach the same conclusion: at this
  campaign's sample size, a learned model of the crowd's own errors is not
  trustworthy, and every probability submitted in this campaign came from the
  hand-built probabilistic pipeline instead.

  Across 86 settled matches and 935 individually scored questions as of this writing,
  the campaign holds a cumulative total of approximately $+3{,}415$ RBP at a 70\%
  crowd-beat rate, moving from the top 5.6\% to the top 1.1\% of an open field of
  nearly 3,900 registered participants over the final three weeks of the group and
  knockout stages. I report the campaign's best and worst individual results in full,
  including an $-80.83$ RBP loss that independently confirms a previously identified,
  still-active failure mode, rather than smoothing over the record.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.align = "center", fig.pos = "H")

required_pkgs <- c("knitr", "kableExtra")
for (pkg in required_pkgs) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg, repos = "https://cloud.r-project.org")
  }
}
library(knitr)
library(kableExtra)
```

\newpage

# Introduction

## Motivation and Problem Framing

The 2026 FIFA World Cup, hosted across the United States, Canada, and Mexico in June
and July 2026, poses an unusual forecasting problem. The object of prediction is not a
continuous outcome---goals, points, margins---but a portfolio of discrete binary and
three-way questions resolved at match level: which team wins, whether each team
scores, whether a named player records a shot on target, how many corners are taken in
the first half. This structure is closer to the question sets on platforms such as
Good Judgment Open and Metaculus than to a standard sports-betting market.

This paper documents an ongoing research campaign I have run inside a specific
competitive environment: the Jump Probability Cup (JTC), a World Cup prediction
tournament on \texttt{play.sportspredict.com} scored by \textit{Relative Brier Points}
(RBP). Under the RBP mechanism, a submitted probability earns positive points when its
squared error is lower than the platform-revealed crowd consensus on the same question
and negative points otherwise. RBP is therefore a \emph{relative} proper scoring rule:
the optimal strategy remains reporting my true belief, yet the metric rewards
calibration relative to the crowd rather than relative to an absolute ground truth. The
competitive advantage lies not in outperforming the crowd on raw accuracy but in
identifying the systematic ways the crowd's aggregate deviates from a calibrated
probability.

Four contributions structure this work. First, I present a complete statistical
forecasting pipeline---Elo ratings, a Poisson goal-rate regression, a Dixon--Coles
correction, Negative-Binomial overdispersion, an ordered logit, and a StatsBomb
event-level panel---grounded in a systematic literature review and validated by
walk-forward backtesting against 49,000-plus international matches. Second, I fit and
periodically revalidate an empirical crowd-bias correction characterising the JTC
crowd's systematic compression toward 50\%. Third, I report four independent tests of
whether a learned model of my own estimate and the crowd's should replace the hand-built
pipeline, all four of which answer no, and I show why that answer is expected rather
than surprising given the campaign's sample size. Fourth, I document a live, resolved
track record across 935 questions spanning the group stage through the quarterfinal,
categorise outcomes by pricing technique, and trace the root causes of both the largest
gains and losses without smoothing over the ones that went badly.

## Scope of This Document

All model code, data files, and per-match research records referenced in this paper
reside in a private git repository (\texttt{sportspredict\_research/}). File paths
cited throughout (e.g., \texttt{model/elo.py}, \texttt{datasets/master\_question\_dataset.csv})
refer to paths within that repository. The master question dataset, at its most recent
rebuild, contains 943 rows across 85 source match files; the most recent settled match,
France versus Morocco (quarterfinal, 2026-07-09), is not yet folded into that rebuild,
and its verified standalone total is reported alongside the master-dataset figure
rather than silently merged into it (\S\ref{sec:results}). Every number in this paper
traces to a specific file read at the time of writing; where two documents in the
repository disagree because they were computed at different points in a still-running
campaign, both figures are reported with the disagreement stated plainly rather than
resolved by assumption.

\newpage

# Literature Review

The modelling choices in this project rest on a systematic review of thirteen sources,
organised by modelling lineage rather than by publication date.

## Goal-Based and Poisson-Family Models

**Dixon \& Coles (1997)** \citep{dixoncoles1997} is the foundational contribution in
this lineage. It treats home and away goals as two independent Poisson distributions:
$$
X \sim \text{Poisson}(\lambda), \quad \lambda = \alpha_i \cdot \beta_j \cdot \gamma
$$
$$
Y \sim \text{Poisson}(\mu), \quad \mu = \alpha_j \cdot \beta_i
$$
where $\alpha_i$ is the attack strength of team $i$, $\beta_i$ is the defensive
weakness, and $\gamma$ is a home-advantage multiplier. In log-linear form this is a
Poisson GLM with team dummy variables---attack and defence---plus a home-effect term.
Two enduring contributions characterise this work: the \emph{tau correction}, a
multiplicative adjustment to the four low-score cells $(0\text{--}0, 1\text{--}0,
0\text{--}1, 1\text{--}1)$ that repairs independent Poisson's tendency to under-predict
draws:
$$
\tau_{\lambda,\mu}(x,y) = \begin{cases}
1 - \lambda\mu\rho & \text{if } x=0, y=0 \\
1 + \lambda\rho    & \text{if } x=0, y=1 \\
1 + \mu\rho        & \text{if } x=1, y=0 \\
1 - \rho           & \text{if } x=1, y=1 \\
1                  & \text{otherwise}
\end{cases}
$$
where $\rho$ is a single dependence parameter estimated from data, and exponential
time-decay weighting, $\phi(t) = e^{-\xi t}$.

**Karlis \& Ntzoufras (2003)** \citep{karlisntzoufras2003} model the home/away goal
correlation directly via a bivariate Poisson (BP) distribution constructed from three
independent latent Poisson components. If $Z_1, Z_2, Z_3$ are independent Poisson with
means $\lambda_1, \lambda_2, \lambda_3$, then $(X_1, X_2) = (Z_1 + Z_3, Z_2 + Z_3) \sim
\text{BP}(\lambda_1, \lambda_2, \lambda_3)$, with $\text{Cov}(X_1, X_2) = \lambda_3 \geq
0$. The base BP model can only capture positive correlation, a structural constraint
relevant to international football, where the empirical data---particularly at the
international level---frequently exhibits negative goal correlation, motivating the
Sarmanov-family extension below.

**Michels, Ötting \& Karlis (2023)** \citep{michels2023} reframe the Dixon--Coles tau
correction as a special case of the Sarmanov family of bivariate distributions and
demonstrate empirically---motivated by women's football, where negative goal
correlation is as strong as $\rho \approx -0.25$ to $-0.40$---that the DC correction
achieves at most $\rho \approx -0.05$ to $-0.08$ at typical scoring rates. This finding
directly bounds the benefit I could expect from re-fitting the $\rho$ parameter in this
project (\S\ref{sec:models}).

**Crowder, Dixon, Ledford \& Robinson (2002)** \citep{crowder2002} extend the DC
framework by allowing attack and defence parameters to evolve as a state-space process
rather than remaining fixed within a season. A full implementation of this formulation
is beyond this project's scope; I substitute DC's exponential time-decay combined with
periodic re-fitting, which approximates time-varying strength without a full particle
filter.

## Rating-Based Models and Outcome Regression

**Hvattum \& Arntzen (2010)** \citep{hvattum2010} furnish the direct template for half
of my modelling stack. The standard Elo update,
$$
R'_h = R_h + K \cdot (S_h - E_h), \quad E_h = \frac{1}{1 + 10^{(R_a - R_h)/400}}
$$
together with a goals-based variant in which the $K$-factor scales with margin of
victory, feeds a \emph{proportional-odds ordered logistic regression} that maps the Elo
difference $x = R_h - R_a$ directly to Home/Draw/Away probabilities:
$$
P(\text{Away win}) = \sigma(\kappa_1 - \beta x), \quad P(\text{Draw}) = \sigma(\kappa_2
- \beta x) - \sigma(\kappa_1 - \beta x)
$$
$$
P(\text{Home win}) = 1 - \sigma(\kappa_2 - \beta x)
$$
where $\sigma(\cdot)$ is the logistic function and $\kappa_1 < \kappa_2$ are fitted
cutpoints. Applied to 14 seasons and 30,524 English league matches, this approach
outperforms uniform and historical-frequency baselines but falls short of
bookmaker-implied probabilities---a result that directly motivates treating market
prices as a primary rather than secondary input wherever they are available.

## Scoring-Rule Verification and Calibration

**Constantinou \& Fenton (2012)** \citep{constantinoufenton2012} introduce the Ranked
Probability Score (RPS) as a scoring rule that respects football's ordinal outcome
structure:
$$
\text{RPS} = \frac{1}{2}\left[(p_A - e_A)^2 + (p_A + p_D - e_A - e_D)^2\right]
$$
where outcomes are ordered Away $<$ Draw $<$ Home on the goal-difference axis. Five
canonical benchmark scenarios demonstrate that the plain Brier score---treating the
three outcomes as nominal---can rank a forecaster who is confidently wrong in direction
no worse than one who is one category off, a pathology RPS avoids by construction.
Whether the JTC platform's weighted-Brier formula inherits this pathology remains an
open empirical question (\S\ref{sec:future}).

**Murphy (1973)** \citep{murphy1973} supplies the foundational calibration--refinement
decomposition of the Brier score,
$$
\mathbb{E}[S(P,X)] = \underbrace{\text{Var}(X)}_{\text{UNC}} -
\underbrace{\mathbb{E}_P\{[\mathbb{E}(X|P) - \mathbb{E}(X)]^2\}}_{\text{RES}} +
\underbrace{\mathbb{E}_P\{[\mathbb{E}(X|P) - P]^2\}}_{\text{REL}}
$$
where RES is resolution (discriminative power) and REL is reliability (calibration
error). This decomposition frames \S\ref{sec:calibration}: it distinguishes "I cannot
separate strong from weak teams" (low RES) from "my confidence is systematically
miscalibrated" (high REL), two failure modes requiring different remedies.

**Foulley (2021)** \citep{foulley2021}, applied to four seasons of UEFA Champions
League group-stage data, yields three findings that bear directly on this project: a
practical estimation procedure for REL, RES, and UNC using isotonic regression, which my
backtesting diagnostics adapt; the result that Draw is structurally the hardest outcome
to discriminate (Harrell's $C \approx 0.62$ for Draw versus $0.79$--$0.82$ for
Home/Away); and a closed-form derivation showing that a Poisson log-linear model with a
shared Elo-difference covariate of opposite sign in each team's predictor automatically
induces negative home/away goal correlation---justification for incorporating the Elo
difference directly into the Poisson means rather than relying solely on the tau
correction.

## Small-Sample Model Selection and Trading-Rule Replication

A fourth thread, absent from the literature review's original scope but load-bearing
for \S\ref{sec:noml}, concerns when a learned estimator should be trusted at all.
\citet{peduzzi1996} and its extension in \citet{riley2019} establish the
events-per-variable (EPV) threshold---roughly 10 to 20 settled outcomes per fitted
parameter---below which logistic-regression coefficients are provably unstable, a
threshold this campaign's own question-level sample sits well under for any
multi-feature blending model. \citet{harveyliuzhu2016} documents that a majority of
published trading-rule discoveries fail to replicate out of sample, the direct analogue
of the named situational rules in \S\ref{sec:rules}, several of which have already
required evidence-forced revision within this campaign's own short history.
\citet{merkle2017} supplies the Brier-score decomposition appropriate to a
heterogeneous question portfolio, used in \S\ref{sec:results}. \citet{mamlouk2023}
shows empirically, in a sports-betting context, that model selection by calibration
rather than raw accuracy is the decision that actually preserves return, the reason
Brier score, not classification accuracy, is the only metric used throughout
\S\ref{sec:noml}.

## Wisdom of Crowds and Probability Aggregation

**Lichtendahl, Winkler \& Bickel (2020)** \citep{wisdomcrowds2018wc} analyse the 2018
FIFA World Cup public forecasting contest (\texttt{fifaexperts.com}), scored by Brier
score. Simple unweighted averaging of participants' probabilities and betting-market
consensus both ranked at or near the top of the leaderboard, outperforming most
individual participants; more sophisticated weighting schemes---including
Budescu--Chen regret-minimisation \citep{budescuchen2015} and online expert-advice
algorithms---did not reliably outperform simple averaging. The paper also proves that
the minimax-optimal forecast under Brier scoring with zero information is the uniform
forecast $(1/3, 1/3, 1/3)$, formal justification for the standing rule
(\S\ref{sec:rules}) against submitting 0\% or 100\% under any circumstances.

This project's core finding---that the JTC's free-entry, financially undisciplined
crowd constitutes a structurally compressed version of a calibrated probability---is
consistent with the favourite--longshot bias literature \citep{snowbergwolfers2010}:
participants without financial exposure systematically avoid extreme probabilities,
causing the aggregate to regress toward 50/50.

\newpage

# Data

## Source Inventory and Reliability Tiers

The most distinctive data artefact in this project is a tiered source-reliability
audit, derived empirically from settled matches spanning the complete group stage and
the knockout phase through the quarterfinal, encompassing over 900 individually graded
questions---not asserted a priori, but constructed by tracing specific wins and losses
to the data source driving each decision. Following \citet{gebru2021}'s
Datasheets-for-Datasets framework, each tier is documented with its intended use, the
basis for its assigned trust level, and known failure modes.

\begin{longtable}{|p{3.5cm}|p{5.8cm}|p{2.9cm}|p{2.9cm}|}
\caption{Data source reliability tiers} \label{tab:datatiers} \\
\hline
\textbf{Source} & \textbf{Use} & \textbf{Trust basis} & \textbf{Failure mode} \\
\hline
\endfirsthead
\multicolumn{4}{c}{\textit{Table \ref{tab:datatiers} continued}} \\
\hline
\textbf{Source} & \textbf{Use} & \textbf{Trust basis} & \textbf{Failure mode} \\
\hline
\endhead
\hline
\multicolumn{4}{r}{\textit{Continued on next page}} \\
\endfoot
\hline
\endlastfoot
\multicolumn{4}{|l|}{\textbf{Tier 1 --- reliably profitable; weight heavily}} \\[4pt]
ESPN roster API & Per-player SOT, fouls, offsides from prior matches (``GD1'' data) &
Largest single wins: $+22$, $+25$, $+20$, $+44$ RBP on offsides/SOT propositions &
Blowout-reversal: team shape abandoned when chasing 3+ goals \\[4pt]
\hline
Main sportsbook consensus (FanDuel + DraftKings + Betfair + bet365 agreeing) & Match
winner, BTTS, major propositions; player \emph{absence} from scorer markets is itself
informative ($p < 0.35$) & When $\geq$3 major books agree within 5 pp, near-ground-truth
& Following a niche aggregator disagreeing with main-book absence: $-11.03$ RBP (Ramos,
POR--UZB) \\[4pt]
\hline
FootyMetrics referee career statistics & Card and penalty questions & Career sample
$\geq$80 matches per referee & Requires blowout discount of $-30$ to $-40\%$ in extreme
mismatches \\[4pt]
\hline
Smarkets exchange (liquid: bid+offer) & Match winner, BTTS, OU goals, OU corners, OU
SOT & Financially-incentivised, two-sided; mid-price is arbitrage-free & Illiquid
(offer-only) markets require a 5.5\% haircut \\[4pt]
\hline
Polymarket ($>$\$1M traded) & Match-winner probability & High liquidity, financially
incentivised & Below \$500K volume, a secondary signal only \\[6pt]
\hline
\multicolumn{4}{|l|}{\textbf{Tier 2 --- useful with specific adjustments; never a
primary driver in isolation}} \\[4pt]
\hline
Footystats 10-match foul averages & Foul-count magnitude prior & Reliable for
magnitude, not for direction & Requires predicting which team will be defending
\\[4pt]
\hline
Model output (Elo / Poisson / ordered logit) & Match structure baseline & Calibrated
against 49K+ historical matches & Weight at $\leq$10\% once Elo gap $> 350$ points;
90\% weight to market \\[6pt]
\hline
\multicolumn{4}{|l|}{\textbf{Tier 3 --- avoid or discount heavily}} \\[4pt]
\hline
Niche aggregator scorer odds (when player absent from main books) & --- & --- &
Goal.com/Ramos failure: $-11.03$ RBP \\[4pt]
\hline
Personal conviction overriding a researched figure & --- & --- & The worst-ever
individual result before the campaign's largest single-question loss (\S\ref{sec:results})
shared this root cause \\
\end{longtable}

The operative source hierarchy, applied whenever sources conflict: verified ESPN GD1
individual behavioural data (with a hard floor of 0.25 on any count-statistic question)
$>$ main sportsbook consensus $>$ referee career statistics $>$ high-volume Smarkets
liquid market $>$ club-level SOT data $>$ Polymarket (above \$1M) $>$ Tier-2 sources $>$
model output in isolation.

## Data Pipeline Architecture

The pipeline follows a staging $\rightarrow$ intermediate $\rightarrow$ mart
architecture in the tradition described by \citet{gebru2021} and visible in open
football-data pipelines such as \texttt{dcaribou/transfermarkt-datasets}. Raw
acquisition is immutable: \texttt{bot/fetch\_markets.py} retrieves all open matches and
markets from the SportsPredict \texttt{v1} API and writes them immediately to disk in
JSONL format, one line per match, without memory buffering prior to write, a convention
I formalised after two early data-loss incidents. A reverse-engineering exercise
documented in \texttt{bot/API\_EXPLORATION\_2026-06-27.md} identified two distinct API
namespaces: a documented external namespace (\texttt{api.sportspredict.com/api/v1},
with two functional routes, \texttt{/matches} and \texttt{/markets}) and an
undocumented browser-only namespace (\texttt{api.sportspredict.com/api},
cookie-authenticated) that additionally exposes settled-trade history, leaderboard
rank, and per-user score summaries.

Classification and cleaning follow: \texttt{bot/classify\_markets.py} categorises every
market question as model-coverable (\texttt{match\_winner},
\texttt{total\_goals\_under/over}, \texttt{btts}) or \texttt{no\_model} (fouls, cards,
corners, offsides, player propositions), with the latter priced via the manual research
and rules workflow. A live Smarkets event-discovery step employs \texttt{limit=300}
plus client-side string filtering, since the \texttt{name=} API query parameter is
confirmed non-functional across all matches---it silently returns an unfiltered page
with HTTP 200, a non-obvious failure mode now standard to check for.

Master dataset construction is the final stage: \texttt{datasets/build\_master\_dataset.py}
reads all per-match files and produces one flat row per (match, question) in
\texttt{datasets/master\_question\_dataset.csv}, the authoritative derived dataset.
Raw files are never edited or overwritten; all reprocessing reads from the original and
writes to a new derived output.

## Master Dataset Documentation

At its most recent rebuild, the master question dataset contains 943 rows across 85
source match files from \texttt{data/external\_markets/} and \texttt{matches/}, of
which 921 carry a settled RBP value and 943 a recorded outcome. Full column
documentation resides in \texttt{datasets/MASTER\_DATASET\_DICTIONARY.md}.

Question-number join keys are unreliable as a matching mechanism: platform numbering
(\texttt{Q1}, \texttt{Q2}, etc.) drifts between the draft submission screen and the
final scoring record, verified concretely in two files
(\texttt{eng\_cro\_2026-06-17.json}: \texttt{q5} pre-match versus \texttt{Q7}
post-match for the same question; \texttt{cro\_pan\_2026-06-23.json}: a one-position
rotation). The build script resolves this via topic-word tokenisation and Jaccard
similarity matching ($\geq 0.35$, requiring $\geq 2$ shared tokens), never by assuming
positional consistency, and the core scored fields are unaffected by the problem since
they derive from the post-match record's own internally consistent numbering. Pre-match
research enrichment covers a genuine minority of rows---roughly a quarter carry a
machine-readable \texttt{recommended\_estimate} from before the match, concentrated in
matches after mid-June, a historical data gap rather than a matching artefact. A small,
explicitly flagged set of questions across the dataset show a submitted value differing
from the researched recommendation by more than 0.03, including the documented
ENG--CRO Kane transcription error discussed in \S\ref{sec:results}. Squad value, age,
and FIFA ranking enrichment is null for five teams (Cape Verde, Curaçao, DR Congo,
Haiti, Ivory Coast), absent from the source Transfermarkt extract rather than a join
failure.

## A StatsBomb Event-Level Panel

A newly constructed StatsBomb open-data panel, restricted to the 2018 and 2022 World
Cups---the subset of the broader 80-competition, roughly 3,961-match public corpus with
schema-identical, fully validated event coverage---gives 128 matches, flattened into a
6,131-row player-match panel and a 257-row team-match panel. I validated this panel
before using it for anything: total goals reproduce known tournament history exactly
(169 goals in 2018, 172 in 2022, 2.641 and 2.688 per match respectively), and top
scorers reproduce correctly (Mbappé's eight goals across the 2022 tournament, including
the final, appear row-for-row).

The panel's value was tested directly rather than assumed. Against a settled match,
Portugal versus Croatia, the pure StatsBomb-history model produced $+93.01$
RBP-equivalent across fourteen comparable questions, while the actual submission for
that match---StatsBomb history blended with live market and tournament-form
evidence---scored $+120.30$ RBP, and the actual submission's advantage held despite the
pure-history model individually winning six of the fourteen questions outright. Every
StatsBomb-derived rate used in this campaign is accordingly treated as a prior blended
with the current tournament's own live form, never a standalone replacement for it.

## Measurement Divergence Across Data Sources

A finding worth stating plainly rather than absorbing silently into the pipeline: ESPN
and StatsBomb do not count the same underlying events identically. Across matches
covered by both sources, ESPN records roughly 20\% fewer fouls, 27\% fewer yellow
cards, 38\% more shots on target, and 30\% more corners than StatsBomb for the same
matches, while total-shot counts are close to parity between the two. The practical
consequence is that raw counts must never be pooled across a \texttt{data\_source}
column without a source-specific correction factor; every count-based feature used
anywhere in this pipeline is tagged by its source and never silently averaged with a
count from the other source.

\newpage

# Methodology \label{sec:models}

I run two parallel approaches to the core match-result question---a goals-based
pipeline and a direct outcome-regression pipeline---following the posture recommended
by \citet{hvattum2010}. For proposition markets outside the reach of either model
(fouls, cards, offsides, individual player propositions), a live Smarkets market price
is the primary input, supplemented by the rules framework of \S\ref{sec:rules}.

## Point-in-Time Elo Ratings (\texttt{model/elo.py})

Every national team carries a single numeric rating updated after each historical match
using the standard Elo update rule. Three domain-specific refinements are applied: a
goal-difference multiplier, $K_{\text{eff}} = K_0 \cdot \frac{11 + |\Delta g|}{8}$,
which imposes diminishing returns to very large victories; a home-advantage offset of a
flat $+100$ rating points applied to the home team when the match is not on a neutral
venue; and a match-importance $K$-factor, $K=60$ for World Cup knockout matches,
decreasing to $K=20$ for friendlies. Ratings are computed over the full
\texttt{martj42} international results panel (49,473 matches, 1872--present), strictly
point-in-time with no lookahead. The initial rating $R_0 = 1500$ applies at a team's
first appearance; by 2026 this matters only for recently FIFA-affiliated nations with
short competitive histories.

**A data defect that silently froze every in-progress rating.** Pulling Elo
differentials for five knockout-stage matches in early July surfaced a bug affecting
every rating used in this paper up to that point. Argentina's pre-match rating is
identical---2189.5282---across all three of its group-stage rows in the project's Elo
panel, despite winning those matches 3--0, 2--0, and 3--1. The cause traces to
\texttt{data/international\_results/results.csv}, the historical panel this project
treats as ground truth: it records every 2026 World Cup match's score as \texttt{NA}, a
fixture list rather than a results file, so \texttt{model/elo.py} has never processed a
single real 2026 result. Every "current" rating used anywhere in this campaign has
actually been frozen at pre-tournament, early-June values. This is not a lookahead
problem---nothing from the future leaked into an earlier estimate---but its opposite: a
model made needlessly uninformative by failing to update on results that were already
known. I built a point-in-time replay script
(\texttt{ml/backtests/r16\_point\_in\_time\_elo\_replay.py}) that reconstructs each
team's actual group-stage and knockout results through the exact Elo formula above,
with true home/neutral status confirmed from venue data rather than assumed. Applied to
five retrospective matches, the replay moved Elo differentials by 12 to 35 points
relative to the frozen baseline (Table \ref{tab:elocorr}). The workaround is match-by-match,
not a pipeline-level fix: it needs to be re-run for every future knockout fixture until
the underlying source file, not just individual matches' numbers, is corrected.

\begin{table}[H]
\centering
\caption{Corrected point-in-time Elo differentials versus the frozen, pre-tournament
baseline, five retrospective matches.} \label{tab:elocorr}
\begin{tabular}{lrrr}
\toprule
Matchup & Corrected & Frozen (stale) & Difference \\
\midrule
Switzerland $-$ Colombia & $-79.45$ & $-111.92$ & $+32.47$ \\
Argentina $-$ Egypt & $+363.62$ & $+379.78$ & $-16.16$ \\
United States $-$ Belgium & $-129.52$ & $-117.22$ & $-12.30$ \\
Portugal $-$ Spain & $-142.62$ & $-169.05$ & $+26.44$ \\
Mexico $-$ England & $-70.36$ & $-105.58$ & $+35.22$ \\
\bottomrule
\end{tabular}
\end{table}

## Poisson Goal-Rate Regression (\texttt{model/poisson\_goals.py})

The Elo difference between two teams is translated into expected goal counts via a
Poisson GLM:
$$
\log(\lambda_{\text{home}}) = b_0 + b_1 \cdot \mathbf{1}[\text{home}] + b_2 \cdot \Delta
R, \quad \log(\lambda_{\text{away}}) = b_0 + b_2 \cdot (-\Delta R)
$$
where $\Delta R = R_{\text{home}} - R_{\text{away}}$ is the pre-match Elo difference and
$\mathbf{1}[\text{home}]$ is the home-advantage indicator, suppressed to zero for
neutral-venue matches. Each historical match contributes two observations, one per
team, exponentially down-weighted by recency with a half-life of approximately 2.4
years. Fitted on $n = 49{,}400$ matches ($n_{\text{obs}} = 98{,}800$ team-match
observations) with exponential recency weighting (decay constant $\xi = 0.0008$
day$^{-1}$), Table \ref{tab:poisson_coefs} reports the MLE estimates.

\begin{table}[H]
\centering
\caption{Fitted Poisson GLM parameters} \label{tab:poisson_coefs}
\begin{tabular}{llrrp{5.5cm}}
\toprule
\textbf{Symbol} & \textbf{Parameter} & \textbf{Estimate} & \textbf{SE} & \textbf{Interpretation} \\
\midrule
$b_0$ & Intercept        & $0.1041$   & $0.0141$              & Log-goals at equal Elo, neutral venue \\
$b_1$ & Home advantage   & $0.2302$   & $0.0211$              & Log-goals uplift at home ($e^{b_1} \approx 1.26\times$ goals) \\
$b_2$ & Elo difference   & $0.001810$ & $3.68 \times 10^{-5}$ & Log-goals uplift per Elo point of advantage \\
$\xi$ & Decay constant   & $0.0008$ day$^{-1}$ & ---          & Exponential recency weight; half-life $\approx 2.4$ yr \\
\bottomrule
\end{tabular}
\end{table}

A second, independent bug was found and resolved in \texttt{build\_design\_matrix()}:
the training code set the home-advantage indicator without consulting the match's
\texttt{neutral} flag, while every downstream file (prediction, backtesting,
rho-fitting) correctly conditions on it, a train/serve mismatch affecting roughly 26\%
of the historical panel's rows. The fix has been applied and the model re-fitted; the
coefficients in Table~\ref{tab:poisson_coefs} reflect the corrected model. Worth
recording directly rather than smoothing over: I had predicted the bug was attenuating
$b_1$ downward from a rough prior of $\sim 0.35$, and the opposite turned out to be
true---the buggy coefficient (0.256) was higher than the fixed one (0.230), because
neutral-venue "home" teams skew slightly toward the higher-rated side by labelling
convention, so including them inflated the home-advantage signal rather than
suppressing it. The fix is correct regardless of which direction the bug ran; the
remaining gap between $b_1=0.230$ and the $\sim 0.35$ prior is better explained by
attenuation bias from a noisy Elo covariate (\S\ref{sec:calibration}) than by this bug.

## Dixon--Coles Correction and Negative-Binomial Overdispersion

The Poisson means above are used to construct a full scoreline probability grid. Two
corrections are applied sequentially. The Dixon--Coles tau correction's dependence
parameter is estimated by maximum-likelihood grid search over the historical panel; the
current fit, $\hat{\rho} = -0.05$, is close to the theoretical maximum
\citet{michels2023} finds achievable at typical scoring rates, so re-fitting via a full
bivariate Poisson would not materially improve calibration. A Negative-Binomial refit
of the marginals addresses a separately documented overdispersion pattern in the
total-goals tail: an overdispersion parameter $\alpha$ is fitted by MLE, with $\alpha =
0$ reducing to the Poisson special case. From the full scoreline grid, every
model-coverable market is computed analytically: match result, total goals over/under,
both-teams-to-score, and "team scores $\geq N$." Following \citet{foulley2021}'s
Appendix B derivation, incorporating the Elo difference directly into the Poisson means
automatically induces negative home/away goal correlation without a separate bivariate-Poisson
$\lambda_3$ term, theoretical grounding for the current model structure.

\begin{table}[H]
\centering
\caption{Fitted Dixon--Coles and Negative-Binomial parameters ($n_{\text{obs}} = 98{,}800$)} \label{tab:dc_nb_coefs}
\begin{tabular}{llrrp{5cm}}
\toprule
\textbf{Symbol} & \textbf{Parameter} & \textbf{Estimate} & \textbf{SE} & \textbf{Interpretation} \\
\midrule
$\hat{\rho}$    & DC low-score dependence & $-0.05$  & ---    & Grid-search MLE; near theoretical achievable maximum \citep{michels2023} \\
$\hat{\alpha}$  & NB dispersion           & $0.0992$ & $0.0131$ & Variance inflation: $\text{Var}(X) = \lambda + \hat{\alpha}\lambda^2$ \\
\bottomrule
\end{tabular}
\footnotetext{SE for $\hat{\rho}$ is not available as it is estimated by grid search rather than gradient-based MLE.}
\end{table}

## Ordered Logit (\texttt{model/ordered\_logit.py})

The direct implementation of \citet{hvattum2010}'s approach fits a proportional-odds
ordered logistic regression against the three-way match outcome using Elo difference
and a home-advantage indicator as covariates, plus two cutpoints. Because it is
calibrated against the exact quantity it predicts, this model backtests with superior
outcome-level calibration relative to the goals-based pipeline in the high-confidence
range. It serves as the primary instrument for match-result markets; the goals-based
pipeline handles all goal-total, BTTS, and proposition questions.

\begin{table}[H]
\centering
\caption{Fitted ordered logit parameters} \label{tab:logit_coefs}
\begin{tabular}{llrrp{5cm}}
\toprule
\textbf{Symbol} & \textbf{Parameter} & \textbf{Estimate} & \textbf{SE} & \textbf{Interpretation} \\
\midrule
$\beta$     & Elo slope           & $0.005199$ & $0.000176$ & Log-odds increase per Elo point of advantage \\
$\delta$    & Home advantage      & $0.3771$   & $0.0745$   & Log-odds boost at non-neutral venue \\
$\kappa_1$  & Lower cutpoint      & $-0.7702$  & $0.0643$   & Away-win / Draw boundary \\
$\kappa_2$  & Upper cutpoint      & $0.5549$   & $0.0634$   & Draw / Home-win boundary \\
\bottomrule
\end{tabular}
\footnotetext{Standard errors computed from the numerical Hessian of the weighted log-likelihood at the MLE; SE for $\kappa_2$ is obtained via the delta method.}
\end{table}

At equal Elo on a neutral venue, the implied probabilities are $P(\text{Away}) =
\sigma(\kappa_1) = 0.316$, $P(\text{Draw}) = 0.319$, $P(\text{Home}) = 0.365$.

## Live Market Pricing via Lambda-Fitting

For questions outside both models' scope, Smarkets market prices serve as the primary
input. A Poisson lambda is recovered from each available over/under threshold via
root-finding,
$$
\hat{\lambda} = \text{brentq}\!\left(f(\lambda) = \text{Poisson.sf}(k, \lambda) - p, \;
(0.001, 100)\right)
$$
where $p$ is the de-vigged market over-probability and $k$ is the threshold. Multiple
thresholds for the same underlying quantity give internal consistency checks: if two
over/under markets on the same statistic imply $\hat{\lambda}$ values differing by more
than $\sim 0.05$, I flag the model as internally inconsistent before submission. For
head-to-head comparison questions, the Skellam distribution yields the exact analytic
probability \citep{gregoryskellam1946}: if $X \sim \text{Pois}(\lambda_A)$ and $Y \sim
\text{Pois}(\lambda_B)$ independently, $P(A > B) = 1 - F_{\text{Skellam}}(0;\,
\lambda_A, \lambda_B)$. For compound questions with no direct market (for example,
"will both teams receive at least one card"), I decompose each side's own
independently measured rate rather than assume a shared or symmetric prior---the single
best individual result in the campaign to date came from exactly this method
(\S\ref{sec:results}).

## Validation: Expanding-Window, Time-Ordered Backtesting

All statistical models are validated via a time-ordered walk-forward backtest
(\texttt{model/backtest\_harness.py}), explicitly not random $k$-fold cross-validation,
since match outcomes are temporally ordered and model coefficients are fitted using
only recency-weighted prior data. The harness reports mean Brier score and mean RPS for
the match-result market against a naïve $\frac{1}{3}/\frac{1}{3}/\frac{1}{3}$ baseline,
together with a per-decile calibration table.

\newpage

# Crowd-Bias Model and Decision Rules \label{sec:rules}

## The Central Empirical Finding

I fit a direct empirical relationship between my submitted estimates and the
platform-revealed crowd consensus across all 772 submitted questions in the campaign
with a paired crowd estimate, spanning the group stage through the quarterfinal:
\begin{equation}
\hat{p}_{\text{crowd}} \approx 0.504 + 0.61 \times (\hat{p}_{\text{mine}} - 0.5)
\end{equation}
Full regression statistics: $r = 0.829$, $R^2 = 0.688$, residual SD $= 7.19$ pp, 95\%
CI on slope $= [0.580,\, 0.638]$. A time-ordered out-of-sample check (first 540
questions chronologically as training, remaining 232 as test) gives RMSE $= 5.85$ pp
and mean bias $= 1.08$ pp.

This relationship is worth reporting alongside an earlier fit rather than only in its
final form: at $n=384$, restricted to the group stage alone, the same regression gave
slope $0.60$, intercept $0.511$, and out-of-sample RMSE $8.44$ pp. Doubling the sample
into the knockout rounds moved the slope by $0.01$ and the intercept by $0.007$, and
tightened the out-of-sample RMSE rather than degrading it. This is the pattern I would
expect if the compression ratio is a genuine platform-level behavioural constant---the
same $\sim$3,900 registered participants, scored under the same rule, with no
financial exposure, generate crowd estimates under an unchanged incentive structure
whether the match is a group-stage fixture or a quarterfinal---rather than an artefact
of an early, unrepresentative sample.

The 0.61 slope indicates the crowd compresses any submitted estimate by roughly 39\% of
the distance back toward 50\%, with a small uniform upward anchor of half a percentage
point. Decomposing by direction: when my estimate falls below 0.5 (500 of 772
questions), my mean (0.334) sits close to the realised base rate, while the crowd's
mean (0.403) overestimates the probability of "unlikely" outcomes by roughly 6.9 pp; in
the above-0.5 direction (252 questions), the crowd's mean (0.588) underestimates
"likely" outcomes by 5.1 pp relative to my own mean (0.639). In aggregate, the crowd is
less extreme than my estimate on 73.2\% of questions and more extreme on only 22.5\%,
direct evidence of systematic compression rather than simple noise around my own
number.

The crowd-bias model does not generate primary estimates; it anticipates the
structural gap between a planned submission and the expected crowd consensus, letting
me quantify in advance the directional magnitude of any proposed deviation. I treat it
as a live, periodically revalidated empirical parameter, updated whenever a new batch
of resolved questions is large enough to shift the regression coefficients by more than
one standard error. A necessary caveat on the performance record: across an early
85-question calibration window, the crowd's raw Brier score in fact exceeded my own,
and my head-to-head beat-crowd rate at that sample point was only 52.9\%. The large
positive cumulative RBP derives from winning disproportionately on higher-leverage
questions where the crowd is most wrong in magnitude, not from superior accuracy in a
simple count sense---a high-variance, positive-expected-value strategy, not a uniformly
reliable edge.

## Named Situational Rules

Beyond the baseline crowd-compression model, fifteen named rules encode situational
corrections derived by tracing specific match outcomes to their proximate cause, each
carrying an explicit evidence grade.

\textit{Rule 1} (match-winner questions) sets the estimate as
$\frac{1}{2}(\hat{p}_{\text{crowd consensus}} + \hat{p}_{\text{verified sharp
market}})$. Across the first three settled match-winner questions, this formula was
never the single worst-performing input of the three available---a minimax-regret
rationale rather than an accuracy claim, given the limited sample it was built on.

\textit{Rule 7} (underdog proposition edge) states that when a team is priced below
$\sim$15\% to win, the edge concentrates in \textit{propositions about the underdog},
not the headline match-winner. First confirmed following Germany 7--1 Curaçao
($+72.47$ RBP), where more than 80\% of the total edge derived from five
Curaçao-specific propositions, and subsequently confirmed at $n \geq 5$.

\textit{Rule 8} (high draw probability: shrink comparison propositions) specifies that
when the pre-match draw probability exceeds $\sim$25\%, any "Team A more [count
statistic] than Team B" estimate is shrunk toward 0.50:
\begin{equation}
p_{\text{final}} = 0.65 \times p_{\text{model}} + 0.35 \times 0.50
\end{equation}
Confirmed on four matches with a net positive result, then refined to a 0.55--0.60 lean
after a counterexample (a one-goal draw with a ten-foul differential) showed the
shrinkage could overcorrect.

\textit{Rule 12} (pre-submission risk gate) requires that before submitting a no-market
comparison proposition, I remove the single most extreme game from either team's ESPN
sample and recompute the estimate. If removal shifts the estimate by more than 10 pp, I
treat it as low-confidence and pull it toward the crowd-formula prediction rather than
retaining full distance from 50\%.

\textit{Rule 15} (extreme-underdog forward suppression) sets the primary attacking
forward's SOT and scoring probability into roughly a 0.20--0.35 band when a team is
priced below $\sim$15--20\% to win. This rule underwent a major evidence-forced
revision when it failed severely ($-31.86$ RBP) on a player whose suppressing evidence
came from a match in which his team was comfortably dominant, not chasing a deficit---the
opposite tactical role. The rule now branches explicitly on game state: a blowout with
the team defending a lead gets suppression; a competitive scoreline with the team
chasing does not, and the suppression may even invert.

A hard submission floor prohibits any submission below 0.25 on count-statistic
questions, adopted after a $-42.56$ RBP result (Haiti 2+ offsides, submitted from a raw
Poisson output near zero, resolved YES after Brazil went 3--0 up and Haiti's defensive
shape collapsed) that a 0.25 floor would have capped at roughly $-10$ to $-15$ RBP.
\textit{Rule 6} prohibits submission of 0 or 100\%, formally motivated by
\citet{wisdomcrowds2018wc}'s minimax proof that the uniform forecast is minimax-optimal
under Brier scoring with zero information, and empirically motivated by the worst
individual result in the campaign's early record, a submission at 1.00 on a
personal-conviction override against a researched figure of 0.58--0.60 that scored the
maximum possible penalty when the match drew.

**A cross-match pattern that generalised and then partly failed.** A synthesis across
five settled matches and 75 questions identified a sharp, near-symmetric pattern in two
specific question shapes. Suppressing a named player's proposition on a genuine
multi-game production drought won 7 of 7 tested instances, net $+144.67$ RBP; blending
a favourite team's own shots-on-target threshold upward from a thin market line toward
its recent empirical average lost 5 of 5, net $-155.64$ RBP. I wrote this up as a
standing rule---default to the market, deviate only downward, never upward, on a
favourite's own SOT threshold---and it was live-tested twice in the following weeks
with opposite results. In Brazil versus Norway, staying at the market on Brazil's own
6-or-more-SOT question rather than blending toward the higher empirical average earned
$+22.57$ RBP, the second-biggest win of that match, and the first live confirmation
that the rule improves real decisions rather than just fitting past data. Days later,
in Canada versus Morocco, the same question shape (Canada's own 4-or-more-SOT
threshold) produced the campaign's worst single result to date: I deviated from the
market toward the empirical average anyway, submitted 0.80 against a market-implied
0.577 and a crowd of 0.46, and Canada did not reach the threshold, costing $-80.83$ RBP.
This was not a failure to know the rule; it was a live test of a theoretically
motivated argument for deliberately deviating past a calibrated market estimate near a
leaderboard-standing consideration (\S\ref{sec:noml}), applied to a signal that was not
in fact high-confidence, and it lost exactly the amount the underlying theory predicts
such a deviation should cost. A separate counter-example keeps the rule itself honest
rather than absolute: in Mexico versus England, staying at the market on England's own
5-or-more-SOT question cost $-16.79$ RBP relative to what a personal-history-grounded
empirical estimate would have won, since the empirical read in that case reflected a
still-relevant behavioural pattern rather than a thin, untested sample. The honest
generalisation across all three matches is narrower than "trust the market over your
own SOT projection": the market specifically beats an upward extrapolation from a thin
team-level sample toward a count threshold, and specifically loses when the deviation
reflects a genuine, regime-tested personal pattern rather than an optimistic team-level
projection.

**Independent per-team decomposition for compound questions.** For any "both sides do
X" question with no direct market, I now default to modelling each side's own rate
independently rather than assuming a shared or symmetric prior. The clearest
validation is Brazil versus Norway's "will both teams receive at least one card"
question: Brazil's own four-game card rate ($\lambda=1.75$) and Norway's own
($\lambda=0.5$) combine to $P(\text{both carded}) = (1-e^{-1.75})(1-e^{-0.5}) = 0.325$,
submitted at 0.33 against a crowd of 0.62; the outcome was NO, for $+59.43$ RBP, the
best single result in the campaign to date. The crowd's 62\% reads as a generic
high-stakes-knockout heuristic applied without checking either team's actual
disciplinary record; pricing off each team's own measured rate instead captured a gap a
shared prior could not see.

The pattern across nearly every rule above is consistent: initial confirmation at small
$n$, followed by a named failure that forces a more conditional, context-sensitive
formulation. Rules confirmed at $n = 1$--$3$ are explicitly designated "watch items"
rather than settled doctrine.

\newpage

# Rejection of Learned Blending Models \label{sec:noml}

The pipeline in \S\ref{sec:models} is a portfolio of hand-derived, individually
theory-grounded rules---an Elo update, a Poisson mean, a named situational adjustment
for a specific player's documented drought. That is exactly the kind of estimator a
learned model is supposed to make obsolete: fit one flexible function to the historical
record of my own estimates, the crowd's estimates, and the resulting outcomes, and let
the data discover the blending weights rather than hand-tuning them match by match. I
built and validated that model, more than once, before rejecting it, and I report the
rejection in full because it is the reason the rest of this paper does not contain a
learned blend anywhere.

## Four Independent Tests

**Platt-scaling recalibration.** A two-parameter logistic fit,
\begin{equation}
\text{logit}(P(\text{outcome}=1)) = a + b \times \text{logit}(\hat{p}_{\text{submitted}})
\end{equation}
run on the live submission record, produces global estimates $\hat{a} = -0.21$, $\hat{b}
= 0.51$---nominally suggesting overconfidence---but the bootstrap 95\% confidence
interval for $b$ comfortably includes $1.0$ (perfect calibration), and a walk-forward
check shows the fitted slope is itself unstable across the split ($b=0.196$
out-of-sample versus $0.510$ in-sample) while applying the correction degrades
out-of-sample Brier by 0.028.

**A learned blend of my estimate and the crowd's.** On a filtered sample of 404
group-stage questions across 42 matches, I fit nineteen candidate features (the two raw
probabilities and their gap, Elo differential, rest-days differential, squad-value
differential, FIFA-ranking differential, pre-match draw probability, and nine binary
indicators for named rules that fired at least ten times in-sample) into two candidate
models: an $L_2$-penalised logistic regression and a shallow histogram
gradient-boosting classifier. Three zero-parameter references---trusting the crowd
alone, trusting my own estimate alone, and a naïve 50/50 average---set the bar either
model had to clear. Table \ref{tab:mlresults} reports both a time-ordered walk-forward
validation and a six-fold grouped cross-validation (grouped by match, so no single
match's questions split across the train/test boundary), against the strongest
baseline available on identical held-out rows.

\begin{table}[H]
\centering
\caption{Learned-model performance against the crowd baseline, two validation schemes.} \label{tab:mlresults}
\begin{tabular}{lrrrrr}
\toprule
Scheme & $n_{\text{test}}$ & Model & Brier & Best baseline & RBP-equiv. \\
\midrule
Walk-forward & 208 & Logistic regression & 0.2545 & 0.2157 & $-817.6$ \\
Walk-forward & 208 & Gradient boosting & 0.2227 & 0.2157 & $-147.9$ \\
Grouped 6-fold & 404 & Logistic regression & 0.2569 & 0.2139 & $-1{,}760.3$ \\
Grouped 6-fold & 404 & Gradient boosting & 0.2315 & 0.2139 & $-721.1$ \\
\bottomrule
\end{tabular}
\end{table}

Both models, under both schemes, are worse than simply trusting the crowd, by a margin
equivalent to hundreds to nearly two thousand relative Brier points over the test
window---an order of magnitude larger than any single-match result in this campaign.
Feature-importance diagnostics on the full sample show every one of the nine named-rule
indicators contributes essentially nothing beyond the continuous features already in
the model: permutation importances are dominated by \texttt{crowd\_estimate} and
\texttt{fifa\_ranking\_diff}, with every rule flag near zero. As \citet{harveyliuzhu2016}
would predict, this does not mean the rules lack genuine out-of-sample edge, only that
a learned model cannot yet separate that edge from noise at this sample size. A
proportionality constant fit directly against 386 previously scored questions with a
known realised RBP gives $\hat{S} \approx 101.3$ RBP per unit of Brier improvement, the
conversion used throughout this section.

**A linear regression of RBP against crowd-deviation.** Run on a broader,
less-filtered sample of 771 rows across 71 matches, a regression of realised RBP on the
gap between my estimate and the crowd's gives an in-sample $R^2$ of only 0.053, with
only a player-prop indicator variable reaching significance; the same model loses to a
naïve baseline out-of-sample under grouped cross-validation (mean-squared error 157.23
versus a baseline of 149.59). A small companion exercise on a single player's own
scoring history, illustrative rather than deployed, makes the mechanism visible: an
in-sample $R^2$ of 0.431 on nine observations looks strong until the corresponding
F-test fails to reach significance ($p=0.677$) and leave-one-out cross-validation shows
the model losing to the player's own naïve career average---exactly the overfitting
pattern the events-per-variable literature predicts at this sample size
\citep{peduzzi1996,riley2019}.

**A closed-form result on the cost of deliberate deviation.** Because RBP is an
additive, per-question, strictly proper scoring rule rather than a convex or
rank-based payoff, deliberately pushing a submission a distance $\delta$ past a
genuinely calibrated estimate has an expected cost of exactly $S \cdot \delta^2$ under
the scale constant above, a guaranteed loss with no offsetting expected benefit; the
Kelly-criterion intuition that justifies aggressive position-sizing under a
multiplicative, compounding payoff does not transfer to an additive, bounded scoring
rule, and the correct frame is closer to a tracking-error budget than a
betting-edge calculation. This is not a purely theoretical caveat: the one live
attempt in this campaign to apply a deliberate, theory-motivated deviation past a
calibrated market estimate (\S\ref{sec:rules}, Canada versus Morocco) cost $-80.83$
RBP, the campaign's worst single result, for exactly the reason the theory predicts---the
underlying signal was not in fact high-confidence, and the quadratic cost of being
wrong about that outweighed any plausible benefit.

## Interpretation

Beating the crowd-alone baseline requires extracting real signal from as many as
nineteen features using two to four hundred rows clustered into a few dozen truly
independent matches. \citet{peduzzi1996} and its extension in \citet{riley2019} put the
safe threshold at 10--20 events per fitted parameter, capping a defensible feature
count at this sample size well below what either candidate model used, and
\citet{mamlouk2023}'s finding that calibration-based model selection reliably beats
accuracy-based selection in sports betting is the reason Brier score, not
classification accuracy, is the only metric used throughout this section. This is not
an implementation failure to be fixed with more careful tuning; it is what "not enough
independent data yet" looks like when measured with an honest out-of-sample test rather
than an in-sample fit. I deploy neither learned model, and I treat the hand-built
probabilistic pipeline of \S\ref{sec:models}---individually theory-grounded,
individually falsifiable, and each rule traceable to a specific won or lost
question---as the higher-information estimator until the settled record is an order of
magnitude larger than it is today. Every probability reported anywhere in this paper
follows from that decision.

\newpage

# Calibration and Validation \label{sec:calibration}

## Historical-Panel Calibration (49,000-Match Backtest)

The walk-forward backtest of the goals-based pipeline surfaces a persistent,
approximately uniform 5--8 pp compression gap: at every predicted $P(\text{home win})$
decile from 0.1 to 0.9, the actual home-win rate exceeds the prediction by a broadly
similar margin throughout the range. Re-fitting the Dixon--Coles $\rho$ parameter
produced negligible movement, ruling out the low-score correction as the primary
driver.

The leading hypothesis, documented in \texttt{data/calibration\_research\_notes.md}, is
attenuation bias from the Elo covariate: because Elo is itself a noisy, partially
bootstrapped estimate of true team strength, its direct inclusion in a regression
attenuates the fitted Elo-difference coefficient toward zero, which mechanically
compresses predicted probabilities toward 50/50, precisely the observed pattern.
Glicko and TrueSkill's explicit treatment of rating uncertainty as a structural
quantity, rather than a point estimate, is a principled alternative I have not
implemented. The recommended next diagnostic step, not yet confirmed run, is to
re-plot the calibration gap in logit space and separately for neutral versus
non-neutral matches, since that split alone discriminates between the attenuation-bias
hypothesis and a home-field-advantage miscalibration hypothesis.

## Live-Record Calibration

The Platt-scaling diagnostic reported in full as the first of four tests in
\S\ref{sec:noml} is, at this sample size, statistically inconclusive rather than a
confirmed finding of overconfidence, and I apply no calibration correction until the
live record grows substantially larger. One structural finding does survive this
caution: questions where my submitted estimate falls in the 70--100\% range earn a mean
RBP of only $+0.65$, compared with $+2.7$ to $+3.1$ in every other estimate band,
consistent with the prediction that when both I and the crowd are already confident,
the exploitable gap between us is minimal.

\newpage

# Track Record and Results \label{sec:results}

## Headline Numbers and a Discrepancy Note

Multiple project documents report cumulative RBP totals at different points in the
campaign's timeline, and I report the disagreement rather than pick one silently.

\begin{table}[H]
\centering
\caption{Cumulative RBP reported across project documents (different snapshots)} \label{tab:rbp}
\begin{tabular}{|p{5.5cm}|r|r|r|p{3cm}|}
\hline
\textbf{Source} & \textbf{Matches} & \textbf{Questions} & \textbf{RBP} & \textbf{As of} \\
\hline
\texttt{master\_question\_dataset.csv} (most recent rebuild) & 85 & 921 & +3{,}246.04 & Most recent rebuild \\
\hline
France vs.\ Morocco (verified, not yet merged) & +1 & +14 & +169.17 & 2026-07-09 \\
\hline
\textbf{Combined, current} & \textbf{86} & \textbf{935} & \textbf{$\approx$+3{,}415.21} & \textbf{This writing} \\
\hline
Earlier record & 49 & 436 & +872.13 & 2026-07-01 \\
\hline
\end{tabular}
\end{table}

The master-dataset figure is treated as authoritative for everything already merged,
since it is derived directly from raw per-match JSON files with an explicit
cross-validation step (\texttt{MASTER\_DATASET\_DICTIONARY.md} independently confirms
a dozen RBP figures against the dataset, all matching exactly). France versus Morocco,
the most recent settled match at time of writing, is reported separately because the
next dataset rebuild has not yet run; I do not fold it into the CSV-derived total by
hand. Beat-crowd rate across the 921 scored questions in the current rebuild is 69.7\%
(642 of 921), up from 58.2\% reported at the July 1 snapshot---an improvement worth
treating cautiously rather than as a clean trend, since it reflects both genuine rule
maturation and a still-accumulating, still-noisy sample.

## Individual Match Results

```{r matchresults, echo=FALSE}
matches <- data.frame(
  Match = c(
    "BRA-NOR (2026-07-05)", "FRA-MAR (2026-07-09)",
    "MEX-ECU (2026-06-30)", "FRA-SWE (2026-06-30)",
    "NED-MAR (2026-06-30)", "ENG-CDR (2026-07-01)",
    "CAN-MAR (2026-07-04)", "JPN-SWE (2026-06-22)",
    "BEL-EGY (2026-06-21)", "GER-CUR (2026-06-14)",
    "CAN-BIH (2026-06-12)"
  ),
  Score = c(
    "Brazil 1-2 Norway", "France 2-0 Morocco",
    "Mexico 2-0", "France 3-0", "1-1 (NED out)", "England 2-1",
    "Canada 0-3 Morocco", "1-1", "1-1 (BEL win)", "Germany 7-1", "1-1"
  ),
  RBP = c(
    "+179.26", "+169.17",
    "+154.20", "+148.01", "+120.30", "+119.36",
    "+58.92", "+70.49", "+63.72", "+72.47", "+42.65"
  ),
  `Beat Crowd` = c(
    "13/15", "13/14",
    "11/15", "14/15", "12/15", "11/15",
    "12/15", "8/10", "9/10", "9/10", "8/10"
  ),
  `Key result` = c(
    "Both-teams-carded compound: +59.43", "Priced with zero market anchor at open",
    "Q12 offsides +49.58", "Direct markets dominate",
    "Brobbey SOT +24.84", "ENG 20+ shots +27.51",
    "Worst single result: Canada 4+ SOT -80.83", "SWE more SOT +23.98", "SOT differential",
    "Underdog props x5", "BTTS props"
  ),
  check.names = FALSE
)
knitr::kable(
  matches,
  booktabs = FALSE,
  longtable = FALSE,
  caption = "Selected individual match results (not an exhaustive listing of all 86 settled matches)",
  align = "llrll",
  format = "latex"
) |>
  kableExtra::kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size = 9
  ) |>
  kableExtra::column_spec(1, border_left = TRUE, border_right = TRUE, width = "3.8cm") |>
  kableExtra::column_spec(2, border_right = TRUE, width = "2.2cm") |>
  kableExtra::column_spec(3, border_right = TRUE, width = "1.4cm") |>
  kableExtra::column_spec(4, border_right = TRUE, width = "1.4cm") |>
  kableExtra::column_spec(5, border_right = TRUE, width = "4.5cm")
```

## Two Matches in Detail

Brazil versus Norway (round of 16, July 5) is documented in full elsewhere in this
project and summarised here because it validates two methodological findings at once.
Norway entered as a 20\% pre-match underdog and won 2--1, and eight of nine pure
direct-market questions beat the crowd by five to fourteen RBP each, the unglamorous
backbone that keeps a match profitable independent of any single standout call. The
match's two largest wins came from the independent per-team decomposition rule
(\S\ref{sec:rules}, $+59.43$ RBP) and from correctly differentiating two superficially
identical "hot streak" player props by whether the streak had been tested against
comparable opposition rather than by raw hit rate alone.

France versus Morocco (quarterfinal, July 9) is the only match in this campaign priced
with no crowd consensus or sharp market available for any of its fifteen questions at
the moment pricing began, the first time in forty-nine prior matches that neither
anchor existed. All fifteen questions were therefore priced by the probabilistic
pipeline of \S\ref{sec:models} alone; a Smarkets quote became available only after the
initial pass, for twelve of the fifteen questions, and was used to revise rather than
originate the estimates. The match settled France 2--0 Morocco for $+169.17$ RBP across
fourteen submitted questions (one question was deliberately withdrawn rather than
submitted on a manufactured justification, discussed in the standalone case study), a
new campaign best at the time it settled.

## Pricing-Technique Category Audit

Across the twenty matches in the dataset carrying an explicit pricing-technique tag---question-category
coverage is sparse in the raw files, so this is a representative sample rather than the
full campaign, and I harmonise two category vocabularies the raw files use across
different points in the project's history rather than treat them as separate
schemes---210 questions classify cleanly into seven techniques (Table
\ref{tab:categories}).

\begin{table}[H]
\centering
\caption{Performance by pricing technique, harmonised sample (210 questions, 20 matches)} \label{tab:categories}
\begin{tabular}{|p{6.5cm}|r|r|r|r|}
\hline
\textbf{Category} & \textbf{N} & \textbf{Total RBP} & \textbf{Avg RBP} & \textbf{Beat-Crowd} \\
\hline
DIRECT (market matches question $\sim$1:1) & 99 & +365.23 & +3.69 & 69.7\% \\
\hline
TEAM\_MODEL (Poisson/Skellam from team markets) & 65 & +234.03 & +3.60 & 72.3\% \\
\hline
PLAYER\_LIQUID (bid+offer player market, near-raw) & 20 & +209.74 & +10.49 & 85.0\% \\
\hline
PLAYER\_ILLIQUID (offer-only, $\times$0.945) & 16 & +116.53 & +7.28 & 68.8\% \\
\hline
BASE\_RATE & 4 & +25.23 & +6.31 & 75.0\% \\
\hline
TIMING\_EARLY (before 1st hydration break) & 5 & +17.21 & +3.44 & 40.0\% \\
\hline
TIMING\_NO\_MARKET (self-built late-window) & 1 & +11.76 & +11.76 & 100\% \\
\hline
\textbf{Total} & \textbf{210} & \textbf{+979.73} & \textbf{+4.67} & \textbf{71.4\%} \\
\hline
\end{tabular}
\end{table}

Two changes from an earlier, smaller audit of this same taxonomy are worth stating
directly rather than updating silently. \texttt{PLAYER\_ILLIQUID} was net negative
($-14.27$ total RBP) in a four-match, sixty-question snapshot taken early in the
knockout stage; it is net positive here, consistent with the 5.5\% illiquid-market
discount rule maturing since that snapshot. \texttt{TIMING\_NO\_MARKET}, that earlier
snapshot's one structurally loss-making category ($-63.72$ total, $-9.10$ average
RBP, 42.9\% beat-crowd), has nearly disappeared from the mix here, appearing in only
one of 210 questions---the direct, intended consequence of a standing rule adopted
specifically because of that finding: shade or avoid late-window, no-market questions
rather than continue self-building an estimate for them. Both changes are evidence a
named rule actually changed behaviour and the behaviour change actually paid off, not
just a bigger sample telling a different story by chance.

## Best and Worst Individual Results

The best single-question result in the campaign is Brazil versus Norway's
both-teams-carded compound question, $+59.43$ RBP, discussed in \S\ref{sec:rules}. The
best single-match result is Brazil versus Norway overall, $+179.26$ RBP.

The worst single-question result in the campaign is Canada versus Morocco's "Canada 4
or more shots on target" question, $-80.83$ RBP, discussed in full in
\S\ref{sec:rules} and \S\ref{sec:noml}. I report this in place of an earlier, smaller
worst-result citation (a $-51.97$ RBP personal-conviction override in an early group-stage
match) because the newer result is both larger in magnitude and a more useful failure
to have on record: it is a traceable methodological failure mode that had already been
named and warned about in writing before it happened again, not a one-off lapse in
process discipline. The earlier result remains the motivation for the standing rule
against ever submitting outside roughly the 0.05--0.95 range, and both losses are kept
in this paper rather than superseding one another, since they illustrate two different
failure modes worth guarding against separately.

## External Validation: Leaderboard Standing

A single match's RBP, even a campaign-best one, is a noisy statistic on its own. A
separate and largely independent check is the platform's global leaderboard,
aggregating every participant's cumulative score across every question submitted all
tournament. Three snapshots, captured over the course of the tournament, are reported
in Table \ref{tab:leaderboard}.

\begin{table}[H]
\centering
\caption{Global leaderboard standing over time (field size and rank independently verified from platform snapshots; July 10 figures self-reported at time of writing).} \label{tab:leaderboard}
\begin{tabular}{lrrrr}
\toprule
Date & Rank & Field size & Percentile & Cumulative score \\
\midrule
2026-06-27 & 200 & 3,564 & Top 5.61\% & 1,059.56 \\
2026-06-29 & 177 & 3,628 & Top 4.88\% & 1,245.36 \\
2026-07-10 & 43 & 3,879 & Top 1.11\% & 3,568.30 \\
\bottomrule
\end{tabular}
\end{table}

The field itself grew over the window (3,564 to 3,879, $+8.8\%$), so percentile rather
than raw rank is the comparable statistic across snapshots: standing moved from the top
5.6\% to the top 1.1\% of an open, global field over roughly two weeks spanning the
transition from the group stage into the knockout rounds. Cumulative score more than
tripled over the same window, consistent with, though not solely attributable to, the
knockout-stage results reported across this campaign's individual matches. I do not
have a verified breakdown of the competing field's composition and make no claim about
it; what is independently verifiable is the field size and the rank within it, and by
that measure the standing reported here places this methodology in the top percentile
of an open field with no entry barrier and several thousand active participants,
moving in the same direction the match-level and question-level results above already
point.

\newpage

# Limitations and Threats to Validity \label{sec:limitations}

Live-record calibration remains genuinely open. As detailed in \S\ref{sec:calibration},
the Platt-scaling diagnostic cannot statistically distinguish genuine overconfidence
from sampling noise at the current sample size. I do not currently know with confidence
whether my live estimates are systematically miscalibrated; if they are, the effect is
neither large enough nor stable enough to justify a correction yet.

The precise functional form of the JTC's Relative Brier Points metric has not been
independently confirmed against the live platform. Whether the platform's scoring
behaves more like plain Brier, which does not respect the ordinal Home/Draw/Away
structure, or like RPS, which does, carries strategic implications: under RPS, hedging
probability mass toward Draw on direction-uncertain match-winner questions is
specifically rewarded, and under plain Brier it is not. This has not been tested
empirically against JTC's actual formula.

The Elo data defect documented in \S\ref{sec:models} remains only partially resolved.
The point-in-time replay script corrects individual matches on demand, but the
underlying source file, \texttt{results.csv}, still records every 2026 result as
missing, so the defect will recur on every future in-progress fixture until the
pipeline is fixed at its source rather than patched match by match.

Three features of the World Cup forecasting setting are not fully captured by any
model in this stack. Teams play at most three group-stage matches before elimination
becomes possible, so the three-match ESPN in-tournament sample and the 128-match
StatsBomb historical panel constitute the entirety of the evidence available for any
given team, unlike the multi-season club-level panels the Dixon-Coles and
Karlis-Ntzoufras lineages were built for. International teams also play infrequently
outside major tournaments and squad composition changes substantially between World
Cups, so a four-year-old Elo rating carries considerably less information than a
four-year-old club rating drawn from a continuous domestic competition would. Finally,
a documented failure mode recurs across multiple matches: applying a team-archetype
lesson learned from one team to a different team that shares only superficial
characteristics, as happened when an "elite teams produce low shot volume" pattern
learned from one South American side was incorrectly transferred to a different team
entirely, producing six below-crowd misses on attacking-volume questions in a single
match.

This remains single-investigator research conducted under competitive time constraints
and evaluated against a partially opaque scoring mechanism, without peer review. All
rules in their present form should be regarded as provisional, in particular those
confirmed on samples of $n = 1$--$3$ observations, and \S\ref{sec:rules}'s own
Canada-Morocco result is a reminder that a rule validated once can still fail badly the
next time it is applied to a signal that only superficially resembles the one that
validated it.

\newpage

# Future Work \label{sec:future}

The most consequential near-term improvement is procedural rather than statistical. A
Kane transcription error, a missed submission deadline, and a corner-mapping near-miss
together account for more than 140 RBP of avoidable loss across the campaign, with no
modelling disagreement in any instance. A final value-by-value comparison between what
is about to be submitted and the researched recommendation, before confirming, would
eliminate this class of error entirely. No model improvement identified in this paper
approaches that magnitude of recoverable loss.

Several diagnostics remain genuinely outstanding rather than merely unfinished. Merging
France versus Morocco, and any subsequent semifinal or final results, into the master
dataset rebuild is the most immediate item, since every headline number in this paper
currently carries a small, explicitly flagged gap while that match sits outside the
rebuilt CSV. Re-running the neutral-versus-non-neutral backtest split in logit space
would confirm whether the Poisson pipeline's residual calibration gap originates in the
now-corrected train/predict mismatch or in errors-in-variables attenuation that would
persist regardless. Extending the point-in-time Elo replay beyond the five
retrospective matches already covered, to every remaining knockout fixture, would
close the gap documented in \S\ref{sec:models} properly rather than match by match.
The live Platt-scaling record has a natural checkpoint once the submission count grows
substantially past its current size, at which point the bootstrap confidence interval
on $b$ in Equation~(3) may finally exclude $b=1.0$ with enough precision to justify a
correction. A per-market-type gradient-boosted model trained on the StatsBomb panel,
already downloaded and flattened into team- and player-match panels, is a concrete next
step for fouls, cards, shots-on-target, corners, and offsides base rates specifically,
validated with the same walk-forward and grouped-by-match discipline that rejected the
crowd-blending models in \S\ref{sec:noml}---a narrower, better-specified modelling
target than "blend my estimate with the crowd's" might clear the events-per-variable
bar this paper's own rejection argument currently fails.

Finally, simulating the five canonical scenarios from \citet{constantinoufenton2012}
against the platform's revealed scores would establish empirically whether JTC's
formula behaves like Brier or RPS, with direct implications for the hedging strategy on
three-way match-winner questions, and stratifying the live record by question
family---score propositions, goal-total propositions, shots on target, timing---would
test whether the miscalibration identified by the global Platt diagnostic concentrates
in specific categories rather than sitting uniformly across all of them.

\newpage

# Conclusion

This paper documents a solo forecasting campaign for the 2026 FIFA World Cup, built on
a formal statistical pipeline and a systematic literature review, and evaluated against
a live, proper-scoring-rule mechanism across 935 individually resolved questions.

Two findings anchor it, and they are worth holding together rather than reading as
separate results. The first is operational: the JTC crowd consensus is systematically
compressed toward 50\% relative to a calibrated probability, a relationship that held
essentially unchanged when the sample doubled from the group stage into the knockout
rounds, and this compression, not any model's absolute accuracy advantage, is the
dominant source of positive RBP. The second is structural and, at this campaign's
sample size, decisive: four independent tests, run at different points and different
sample sizes, all reach the conclusion that a learned blend of my own estimate and the
crowd's is not trustworthy, and a live attempt to override that conclusion once, under
a theoretically motivated argument for deliberate deviation, produced the campaign's
single worst result at almost exactly the cost the underlying theory predicts. Neither
finding is a statement that learned models are wrong in general; both are quantified
statements about where the events-per-variable constraint actually binds for this
specific problem at this specific sample size, and both are the reason every
probability in this paper came from a hand-built, individually falsifiable rule rather
than a fitted blend.

The pricing-technique audit gives a concrete, evidence-graded account of which
techniques generate positive expected value: direct market trust and team-level
decomposition are reliably profitable, and a previously identified loss-making
category, self-built late-window timing with no market reference, has nearly vanished
from the mix as the standing rule built to avoid it actually changed behaviour. The
leaderboard trajectory---from the top 5.6\% to the top 1.1\% of an open field of nearly
3,900 participants over three weeks---is the most externally verifiable evidence this
paper offers that these findings, in aggregate, translate into a real result rather
than a favourable read of my own bookkeeping.

\newpage

# Appendix: Data Dictionary Summary \label{sec:appendix}

## \texttt{master\_question\_dataset.csv} Column Inventory (Selected)

Full documentation in \texttt{datasets/MASTER\_DATASET\_DICTIONARY.md}. Built by
\texttt{datasets/build\_master\_dataset.py} from raw per-match files in
\texttt{data/external\_markets/} and \texttt{matches/}. 943 rows, 107 columns, 85
matches at the most recent rebuild.

\begin{longtable}{|p{4.8cm}|p{10cm}|}
\caption{\texttt{master\_question\_dataset.csv}: key columns} \label{tab:dictionary} \\
\hline
\textbf{Column} & \textbf{Description} \\
\hline
\endfirsthead
\multicolumn{2}{c}{\textit{Table \ref{tab:dictionary} continued}} \\
\hline
\textbf{Column} & \textbf{Description} \\
\hline
\endhead
\hline \multicolumn{2}{r}{\textit{Continued on next page}} \\ \endfoot
\hline \endlastfoot
\multicolumn{2}{|l|}{\textit{Match-level (repeated across each match's rows)}} \\[3pt]
\hline
\texttt{match}             & Canonical team-code pair, e.g.\ \texttt{ARG-AUT} \\
\hline
\texttt{date}              & Match date \\
\hline
\texttt{schema\_era}       & 1, 2, or 3 (generation of raw-file schema; each handled separately) \\
\hline
\texttt{final\_score}      & Reported final score \\
\hline
\texttt{match\_total\_rbp} & Summed RBP across all questions for this match \\
\hline
\texttt{elo\_team\_a\_pre}, \texttt{elo\_team\_b\_pre}, \texttt{elo\_diff} & Pre-match Elo ratings and difference \\
\hline
\texttt{rest\_days\_a}, \texttt{rest\_days\_b} & Days since each team's previous match \\
\hline
\texttt{squad\_value\_a\_eur}, \texttt{squad\_value\_b\_eur} & Transfermarkt squad values (EUR; null for 5 teams) \\
\hline
\texttt{draw\_probability} & Pre-match three-way draw probability from verified market \\
\hline
\texttt{rule1\_fired} \ldots \texttt{rule15\_fired} & Boolean flags for each named rule application \\[6pt]
\hline
\multicolumn{2}{|l|}{\textit{Question-level (one value per row)}} \\[3pt]
\hline
\texttt{question\_text}    & Platform question wording \\
\hline
\texttt{question\_category} & Pricing technique: \texttt{DIRECT}, \texttt{TEAM\_MODEL}, \texttt{PLAYER\_LIQUID}, \texttt{PLAYER\_ILLIQUID}, etc.; sparse coverage (210 of 943 rows) \\
\hline
\texttt{recommended\_estimate} & Pre-match research figure; matched by topic-word Jaccard similarity ($\geq 0.35$) \\
\hline
\texttt{our\_estimate}     & Final submitted probability \\
\hline
\texttt{crowd\_estimate}   & Platform-revealed crowd consensus after resolution \\
\hline
\texttt{outcome}           & Binary outcome (1 = event occurred) \\
\hline
\texttt{rbp}               & Relative Brier Points earned (blank, not zero, for unsubmitted questions) \\
\hline
\texttt{beat\_crowd}       & Boolean: did the submitted estimate outperform the crowd? \\
\hline
\texttt{submission\_error\_flag} & True where $|p_{\text{submitted}} - p_{\text{recommended}}| > 0.03$ \\
\hline
\texttt{actually\_submitted} & False for questions never reaching the platform before a deadline, or belonging to a calibration-only match \\
\hline
\texttt{postmortem\_note}  & Free-text root-cause annotation \\
\end{longtable}

## Other Datasets Referenced in This Paper

\begin{table}[H]
\centering
\caption{Supporting datasets} \label{tab:datasets}
\begin{tabular}{|p{5.8cm}|r|r|p{5.2cm}|}
\hline
\textbf{File} & \textbf{Rows} & \textbf{Cols} & \textbf{Role} \\
\hline
\texttt{data/processed/elo\_match\_panel.csv} & $\sim$49,400 & --- & Point-in-time Elo for both teams, one row per historical international match (1872--present) \\
\hline
\texttt{data/international\_results/results.csv} & 49,473 & 9 & Source match results panel (\texttt{martj42}); foundation for all statistical modelling; records every 2026 World Cup score as missing (\S\ref{sec:models}) \\
\hline
StatsBomb WC2018+WC2022 event panel & 6,131 (player-match) / 257 (team-match) & --- & Validated historical event-level panel, 128 matches (\S\ref{sec:models}) \\
\hline
\texttt{ml/feature\_matrix.csv} & 264 & 11 & Meta-model input table, 2026-06-11 to 2026-06-30 \\
\hline
\end{tabular}
\end{table}

\newpage

# References
